---
title: "ESM 244 Assignment 1 - Task 1"
author: "Indigo Bannister"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# set up libraries
library(tidyverse)
library(here)
library(readxl)
library(lubridate)
library(patchwork)
library(ggthemes)

```
# Task 1: Data wrangling & visualization of Sierra amphibians

### Step 1: Call in data

```{r}
# read in xlsx of data
RAMU <- read_xlsx(here("data","sierra_amphibians.xlsx")) %>% 
  filter(amphibian_species == 'RAMU') %>% 
  mutate(survey_date = ymd(survey_date)) %>% 
  mutate(survey_year = year(survey_date)) 

```

### Step 2: Create a graph of count of Rana muscosa by year, grouped by life stage
```{r}
# Filter, clean, and summarize data for analysis
year_lifestage <- RAMU %>% 
  filter(amphibian_life_stage != 'EggMass') %>% 
  group_by(amphibian_life_stage, survey_year) %>% 
  summarize(count = sum(amphibian_number))

# create graph 
graph_1 <- ggplot(data = year_lifestage, aes(x = survey_year, y = count, fill = amphibian_life_stage)) + 
  geom_col() + 
  labs(x = "Year",
       y = "Number of frogs",
       fill = "Life Stage") +
    scale_x_continuous(expand = c(0, 0),
                     breaks = seq(1995, 2002, by = 1)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() + 
  coord_flip() +
  scale_fill_tableau(palette = "Miller Stone", direction = -1) + 
  theme(legend.position = c(0.8, 0.2))

```

### Step 3: Create a graph of frog counts observed in the top 5 lakes
```{r}
# Wrangle data for graph: filter initial dataset, select top 5 counts, and adjust naming for graph visual
top5_lakes <- RAMU %>% 
  filter(amphibian_life_stage != 'EggMass') %>% 
  filter(amphibian_life_stage != 'Tadpole') %>% 
  group_by(lake_id) %>% 
  summarize(count = sum(amphibian_number)) %>% 
  slice_max(order_by = count, n = 5) %>% 
  mutate(lake_name = paste('Lake #',lake_id, sep = ' ')) %>% 
  mutate(lake_name = fct_reorder(lake_name, count))

# create graph of frog counts observed in top 5 lakes
graph_2 <- ggplot(data = top5_lakes, aes(x = lake_name, y = count, fill = lake_name)) + 
  geom_col() + 
  labs(x = "Lake ID",
       y = "Number of frogs")  +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() + 
  coord_flip() +
  scale_fill_tableau(palette = "Blue-Teal", type = "ordered-sequential", direction = 1) +
  theme(legend.position = "none")
  
```

### Step 4: Display both graphs together using patchwork()

```{r}
# Display graphs side by side and add figure caption
graph_1 | graph_2
```

<font size="2"> ***Figure 1:*** *Mountain Yellow-Legged Frog (Rana muscosa) counts by year (separated by life stage) [right] and by lake for adults and subadults [left].* </font>


